Parallel air palletization and tour planning for simultaneous pickup and delivery

### 2023-03-25 latex/comp_or.tex
Paper CompOR update

We also contributed by carrying out multiple experiments with the multi-processing heuristics on synthetic data based on real data from the {\it Brazilian Air Force} transportation history and a new procedure to minimize the distance from the next node destined pallet to the ramp door.

@@

{\color{blue} $k$ is the node index and $\pi(k)$ is the node in position $k$ of tour $\pi$. $\tau^{\pi(k)}$ is the total torque applied by the loaded pallets to the aircraft in this node. }

Let $M^{\pi(k)} = \{p_1, p_2, \ldots, p_m \}$ the set of $m$ pallets. Each pallet $p_i$, $1 \leq i \leq m$, has weight capacity $p_i.w$, volume capacity $p_i.v$, pallet destinations $p_i.to^{\pi(k)}$, $0 \leq k \leq K$, and distance to the CG of aircraft $p_i.d$. $p_i.to^{\pi(k)}$ denotes that pallet $p_i$ may assume a different destination in each node ${\pi(k)}$.

Let $N^{\pi(k)} = \{t^{\pi(k)}_1, t^{\pi(k)}_2, \ldots, t^{\pi(k)}_{n^{\pi(k)}} \}$ be the set of $n^{\pi(k)}$ items to be loaded in node $l^{\pi(k)}$, $0 \leq k \leq K$. Each item $t^{\pi(k)}_j$, $1 \leq j \leq n^{\pi(k)}$, has score $t^{\pi(k)}_j.s$, weight $t^{\pi(k)}_j.w$, volume $t^{\pi(k)}_j.v$, and destination $t^{\pi(k)}_j.to \in L^{\pi(k)}$. Let $N = \bigcup_{0 \leq k \leq K} N^{\pi(k)}$ be the set of items of all nodes along a tour.

Let $Q^{\pi(k)} = \{a^{\pi(k)}_1, a^{\pi(k)}_2, \ldots, a^{\pi(k)}_{m} \}$ be the set of consolidated items loaded in $m_k \leq m$ pallets when the aircraft arrives at node $l^{\pi(k)}$, with $0 \leq k \leq K$. $a^{\pi(k)}_i$, $1 \leq i \leq m$, is the group of picked-up items that were allocated on pallet $p_i$ in some of the previous nodes. $a^{\pi(k)}_i$ has total weight $a^{\pi(k)}_i.w$, total volume $a^{\pi(k)}_i.v$, and destination $a^{\pi(k)}_i.to \in L^{\pi(k)} \cup \{l^{\pi(k)}\}$. If $a^{\pi(k)}_i.to = l^{\pi(k)}$, then $a^{\pi(k)}_i$ is unloaded, and $p_i$ will be available for reloading; otherwise, $a^{\pi(k)}_i$ remains on the aircraft, eventually in another pallet, and with items of $N^{\pi(k)}$ having the same destination.

Let $X_{ij}^{\pi(k)}$ and $Y_{iq}^{\pi(k)}$ be binary variables, where $0 \leq k \leq K$, $1 \leq j \leq n^{\pi(k)}$, $1 \leq i \leq m$ and $1 \leq q \leq m$. $X_{ij}^{\pi(k)} = 1$ if $t_j^{\pi(k)}$ is assigned to $p_i$ in node $l^{\pi(k)}$, and 0 otherwise.

$Y_{iq}^{\pi(k)} = 1$ if $a_q^{\pi(k)}$ is assigned to $p_i$ in node $l^{\pi(k)}$, and 0 otherwise. By definition, $Y_{iq}^0=0$.

Allocations of items or consolidated to pallets in node $l^{\pi(k)}$ can be seen as a bipartite graph $G^{\pi(k)}(V^{\pi(k)}, E^{\pi(k)})$,
where $V^{\pi(k)} = M^{\pi(k)} \cup N^{\pi(k)} \cup Q^{\pi(k)}$, $E^{\pi(k)} = E_N^{\pi(k)} \cup E_Q^{\pi(k)}$,
$(p_i, t_j^{\pi(k)}) \in E_N^{\pi(k)}$
if $X_{ij}^{\pi(k)} = 1$, and $(p_i, a_q^{\pi(k)}) \in E_Q^{\pi(k)}$ if $Y_{iq}^{\pi(k)} = 1$.

@@

\begin{equation} \label{eq:scores}
	\tilde{s} = \sum_{k=0}^{K} \sum_{i=1}^{m} \sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.s
\end{equation}

\begin{equation} \label{eq:costs}
	\tilde{c} = c_{0,\pi(1)}\times(1+c_g\times|\epsilon_0|) + \sum_{k=1}^{K-1} [ c^{\pi(k), \pi(k+1)}\times(1+c_g\times|\epsilon^{\pi(k)}|) ] + c^{\pi(K),0}\times(1+c_g\times|\epsilon_K|)
\end{equation}

@@

\begin{equation} \label{eq:tau}
\tau^{\pi(k)} = \sum_{i=1}^{m}[ p_i.d \times (\sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w +  \sum_{q=1}^{m} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w)];\ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:eps}
\epsilon^{\pi(k)} = \frac{\tau^{\pi(k)}}{maxW \times limit^{CG}_{long}};\ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp11}
	L_0 = L; L^{\pi(k)} = L^{\pi(k-1)} - \{l^{\pi(k)}\}; \ k \in \{1, 2, \ldots, K\}
\end{equation}

@@

\begin{equation} \label{eq:pdp12}
	X_{ij}^{\pi(k)} = 0 \ \mbox{if} \ t_j^{\pi(k)}.to \notin L^{\pi(k)}; \ i \in \{1, 2, \ldots, m\}; \ j \in \{1, 2, \ldots, n^{\pi(k)}\}; \ k \in \{1,2, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp9}
	Y_{iq}^{\pi(k)} = 0 \ \mbox{if} \ a_i^{\pi(k)}.to \notin L^{\pi(k)}; \ i \in \{1, 2, \ldots, m\}; \ q \in \{1, 2, \ldots, m\}; \ k \in \{ 1, 2, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:cons2}
	a_i^{\pi(k+1)}.w = \sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w + \sum_{q=1}^{m} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w;  \ i \in \{1, 2, \ldots, m\}; k \in \{0, 1, \ldots, K-1\}
\end{equation}

\begin{equation} \label{eq:cons3}
	a_i^{\pi(k+1)}.v = \sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.v + \sum_{q=1}^{m_k} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.v;  \ i \in \{1, 2, \ldots, m\}; k \in \{0, 1, \ldots, K-1\}
\end{equation}

\begin{equation} \label{eq:cons5}
	a_i^{\pi(k+1)}.to = t^{\pi(k)}_j.to  \ \mbox{if} \ X^{\pi(k)}_{ij} = 1  \ \mbox{and} \ t^{\pi(k)}_j.to \in L_{\pi(k+1)};  \ i \in \{1, 2, \ldots, m\}; \ j \in \{1, 2, \ldots, n^{\pi(k)}\}; \ k \in \{0,1, \ldots, K-1\}
\end{equation}

\begin{equation} \label{eq:LatIt}
	LatIt^{\pi(k)} = \sum_{i=1}^{m} \sum_{j=1}^{n^{\pi(k)}} ( X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w \times (i\%2) - X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w \times (i+1)\%2 ); \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:LatCons}
	LatCons^{\pi(k)} =  \sum_{i=1}^{m} \sum_{q=1}^{m}  ( Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w \times (i\%2) - Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w \times (i+1)\%2); \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:torqlat}
	s.t.: d^{CG}_{pallet} \times | LatIt^{\pi(k)} + LatCons^{\pi(k)} | \leq  \sum_{i=1}^{m}p_i.w \times limit^{CG}_{lat}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:torqlong}
	s.t.: |\tau^{\pi(k)}| \leq maxW \times limit^{CG}_{long};\ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:payload}
	s.t.: \sum_{i=1}^{m} (\sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w + \sum_{q=1}^{m_k} Y_{iq}^{\pi(k)} \times a_q^k.w ) \leq maxW; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app2}
	s.t.: \sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.w + \sum_{q=1}^{m_k} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.w  \leq p_i.w; \ i \in \{1, 2, \ldots, m\}; \ {\pi(k)} \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app3}
	s.t.: \sum_{j=1}^{n^{\pi(k)}} X_{ij}^{\pi(k)} \times t_j^{\pi(k)}.v + \sum_{q=1}^{m} Y_{iq}^{\pi(k)} \times a_q^{\pi(k)}.v  \leq\ p_i.v; \ i \in \{1, 2, \ldots, m\}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app4}
	s.t.: \sum_{i=1}^{m} X_{ij}^{\pi(k)} \leq 1; \ j \in \{1, 2, \ldots, n^{\pi(k)}\}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app5}
	s.t.:  Y_{iq}^{\pi(k)} = 1 \ \mbox{if} \ a^{\pi(k)}_q.to \in L^{\pi(k)}; \ q \in \{1, 2, \ldots, m\}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp8}
	s.t.: p_i.to^{\pi(k)} = t^{\pi(k)}_j.to\ \mbox{if} \ X_{ij}^{\pi(k)} = 1; \ i \in \{1, 2, \ldots, m\}; \ j \in \{1, 2, \ldots, n^{\pi(k)}\}; \ k \in \{1, 2, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp2}
	s.t.:  p_i.to^{\pi(k)} = a^{\pi(k)}_q.to\ \mbox{if} \ Y_{iq}^{\pi(k)} = 1; \ i \in \{1, 2, \ldots, m\};\ q \in \{1, 2, \ldots, m\}; \ k \in \{1, 2, \ldots, K\}
\end{equation}

The objective of this problem is to find a permutation $\pi \in S_K$\/ that maximizes the function $f_\pi(\tilde{s},\tilde{c})$ \ref{eq:maxf}. In this way, the flight plan will be $l_0, l_{\pi(1)}, \ldots, l^{\pi(K)}, l_0$. $\tilde{s}$\/ is the total score of transported items \ref{eq:scores} and  $\tilde{c}$\/ is the total cost of fuel consumed \ref{eq:costs}. As can be seen, $\tilde{c}$\/ corresponds to the fuel consumption due to the flights carried out and the CG deviation of the transported cargo. Throughout this work, for simplicity, we use $f=\tilde{s}/\tilde{c}$.

@@

The tactic will be, at each shipping node, to predefine the destinations of the pallets at that node. In this way, we will reserve a number of pallets proportional to the volume demanded by each destination at the shipping node. We could have used another criterion, but it was observed in the experiments that volume is more constrictive in airlift. Once the destinations of the pallets are defined, we will use serial and Multi-processing heuristics to find the best possible node-by-node solutions. This strategy is summarized in Algorithm \ref{alg:main}, retrieved from \cite{MesquitaSanches2023}.

@@

The other parameter $volume$\/ is a value greater than 1, which corresponds, at each node $l^{\pi(k)}$, to the ratio between the sum of the volumes of the items ($\sum_{j=1}^{n^{\pi(k)}} t^{\pi(k)}_j.v$) and the load capacity of the pallets ($\sum_{i=1}^{m} p_i.v$). This parameter is passed to $ItemsGeneration$ (line \ref{main:items}), responsible for creating the items to be shipped, which will be presented in the next section.

@@

As we mentioned in the previous section, all tours start and end at $l_0$\/ (lines \ref{tour:pi1}-\ref{tour:pi2}).
After initializing the score and cost values (lines \ref{tour:score}-\ref{tour:cost}), there is a loop for the $K+1$\/ flights (lines \ref{tour:loop1}-\ref{tour:loop2}). Initially we set pallets destination as $-1$\/ (line \ref{tour:-1}). When the aircraft is at node $l_0$, the initial graph $G_1$\/ is empty because it has no consolidated items \ref{tour:g11}. Otherwise, the set $L_k$\/ of remaining nodes when departing from $\pi(k)$ is updated (line \ref{tour:lk1}), and $UpdateConsolidated$\/ (line \ref{tour:dest}) returns the set of consolidated items that have not yet reached their destinations and remain on board, rearranging them on the pallets to minimize CG deviation. This allocation is stored in graph $G_1$\/ (line \ref{tour:g12}).

@@

		\State $\pi(0) \gets 0$ \label{tour:pi1}
		\State $\pi(K+1) \gets 0$ \label{tour:pi2}
		\State $score \gets 0$ \label{tour:score}
		\State $cost \gets 0$ \label{tour:cost}
		\For {$k\gets0$ to $K$} \label{tour:loop1}		
			\For{$i \gets 1$ to $m$}
				\State $p_i.to[k] \gets -1$ \label{tour:-1} \Comment{reset this pallet destination}
			\EndFor	
			\If {$k = 0$}
				\State $L_0 \gets L$
				\State Let $G_1(M \cup N_0, \varnothing)$ \label{tour:g11}
			\Else
				\State $L_k \gets L_k - \pi(k)$  \label{tour:lk1}			
				\State $Q^{\pi(k)}, E_Q^{\pi(k)}, M^{\pi(k)} \gets UpdateConsolidated(\pi(k))$ \label{tour:dest}			
				\State Let $G_1(M^{\pi(k)} \cup N^{\pi(k)} \cup Q^{\pi(k)}, E_Q^{\pi(k)})$ \label{tour:g12}
			\EndIf  \label{tour:lk2}	
			\State $M^{\pi(k)} \gets SetPalletsDestination( \pi(k) )$ \label{tour:dest2}		
			\State $G_2 \gets SolveNode(method, \pi(k), G_1, timeLim)$ \label{tour:node}

@@

Then $MinCGDeviation$ is run through a MIP solver to relocate the consolidated items on the pallets minimizing torque and ensuring that they all remain on board, one on each pallet. As there are few variables, the MIP solver returns an allocation $E^Q_k$ very quickly.

@@

In this subsection we present implementations of $SolveNode(...)$, where $method$\/ corresponds to a heuristic, $k$\/ is the index of the current node $l^{\pi(k)}$ and $G$\/ is the allocation graph of the consolidated items that remain on board at $l^{\pi(k)}$.

@@

The selection of edges for $E_N^{\pi(k)}$\/ uses the {\it edge attractiveness}\/ $\theta^{\pi(k)}_{ij}$, Equation \ref{eq:edge}, which can be understood as the tendency to allocate the item $t^{\pi(k)}_j$\/ to the pallet $p_i$. It is directly proportional to the score, and inversely relative to the volume and torque of each item.

\begin{equation} \label{eq:edge}
	\theta^{\pi(k)}_{ij}= \frac{t^{\pi(k)}_j.s}{t^{\pi(k)}_j.v \times 3000}\times(1-\frac{t^{\pi(k)}_j.w\times|p_i.d|}{\max_s\{t^{\pi(k)}_s.w\}\times\max_q\{|p_q.d|\}});\ i \in \{1,2,\ldots,m\},\ j \in \{1,2,\ldots,n_k\}
\end{equation}

@@

In the pseudo-codes presented below, given an allocation graph $G$\/ for the node $l^{\pi(k)}$, $f_s(G)$\/ is equivalent to the score $s$\/ returned by $ScoreAndDeviation({\pi(k)},G) $, described in \cite{MesquitaSanches2023}.

@@

According to \cite[p.226]{manfrin2006}, there are five topologies for multi-process interchange of information: {\it fully-connected}, where the master process broadcasts to all remaining child processes; {\it replace-worst}, where the best-so-far solution process broadcasts only to the current worst solution process; {\it hypercube}, where processes are connected as a hypercube, and a vertex process broadcasts only to the connected vertices; {\it ring}, in which one process only sends a message to the next process connected to it; and {\it parallel independent runs (PIR)}, in which there are no communication costs and the best solution is chosen among all processes. They concluded that the PIR topology presented, in average, the smallest distance from the optimal solutions.

@@

$3^{st}$ - {\it Shims} set construction and best {\it Shims} selection

All executed in a multi-process run-time, accessing concurrently, reading and writing the list of items and the current value of the CG deviation.
As to the multi-process run-time, we considered each pallet as a parallel process, taking advantage of nowadays multi-core computers.

@@

		\Procedure{$mpShims$}{${N^{\pi(k)}, \pi(k)}, limit, G$}

		\State Let $G(M^{\pi(k)},\ Q^{\pi(k)},\ E^{\pi(k)})$ \label{mpShims:g1} \Comment{the initial solution}

>>>>

We also contributed by carrying out multiple experiments with the multiprocessing heuristics on synthetic data based on real data from the {\it Brazilian Air Force} transportation history and a new procedure to minimize the distance from the next node destined pallet to the ramp door.

@@

{\color{blue} $k$ is the node index and $\pi_k$ is the node in position $k$ of tour $\pi$. $\tau^{\pi_k}$ is the total torque applied by the loaded pallets to the aircraft in this node. }

Let $M^{\pi_k} = \{p_1, p_2, \ldots, p_m \}$ the set of $m$ pallets. Each pallet $p_i$, $1 \leq i \leq m$, has weight capacity $p_i.w$, volume capacity $p_i.v$, pallet destinations $p_i.to^{\pi_k}$, $0 \leq k \leq K$, and distance to the CG of aircraft $p_i.d$. $p_i.to^{\pi_k}$ denotes that pallet $p_i$ may assume a different destination in each node ${\pi_k}$.

Let $N^{\pi_k} = \{t^{\pi_k}_1, t^{\pi_k}_2, \ldots, t^{\pi_k}_{n^{\pi_k}} \}$ be the set of $n^{\pi_k}$ items to be loaded in node $\pi_k$, $0 \leq k \leq K$. Each item $t^{\pi_k}_j$, $1 \leq j \leq n^{\pi_k}$, has score $t^{\pi_k}_j.s$, weight $t^{\pi_k}_j.w$, volume $t^{\pi_k}_j.v$, and destination $t^{\pi_k}_j.to \in L^{\pi_k}$. Let $N = \bigcup_{0 \leq k \leq K} N^{\pi_k}$ be the set of items of all nodes along a tour.

Let $Q^{\pi_k} = \{a^{\pi_k}_1, a^{\pi_k}_2, \ldots, a^{\pi_k}_{m} \}$ be the set of consolidated items loaded in $m_k \leq m$ pallets when the aircraft arrives at node $\pi_k$, with $0 \leq k \leq K$. $a^{\pi_k}_i$, $1 \leq i \leq m$, is the group of picked-up items that were allocated on pallet $p_i$ in some of the previous nodes. $a^{\pi_k}_i$ has total weight $a^{\pi_k}_i.w$, total volume $a^{\pi_k}_i.v$, and destination $a^{\pi_k}_i.to \in L^{\pi_k} \cup \{l^{\pi_k}\}$. If $a^{\pi_k}_i.to = l^{\pi_k}$, then $a^{\pi_k}_i$ is unloaded, and $p_i$ will be available for reloading; otherwise, $a^{\pi_k}_i$ remains on the aircraft, eventually in another pallet, and with items of $N^{\pi_k}$ having the same destination.

Let $X_{ij}^{\pi_k}$ and $Y_{iq}^{\pi_k}$ be binary variables, where $0 \leq k \leq K$, $1 \leq j \leq n^{\pi_k}$, $1 \leq i \leq m$ and $1 \leq q \leq m$. $X_{ij}^{\pi_k} = 1$ if $t_j^{\pi_k}$ is assigned to $p_i$ in node $\pi_k$, and 0 otherwise.

$Y_{iq}^{\pi_k} = 1$ if $a_q^{\pi_k}$ is assigned to $p_i$ in node $\pi_k$, and 0 otherwise. By definition, $Y_{iq}^0=0$.

Allocations of items or consolidated to pallets in node $\pi_k$ can be seen as a bipartite graph $G^{\pi_k}(V^{\pi_k}, E^{\pi_k})$,
where $V^{\pi_k} = M^{\pi_k} \cup N^{\pi_k} \cup Q^{\pi_k}$, $E^{\pi_k} = E_N^{\pi_k} \cup E_Q^{\pi_k}$,
$(p_i, t_j^{\pi_k}) \in E_N^{\pi_k}$
if $X_{ij}^{\pi_k} = 1$, and $(p_i, a_q^{\pi_k}) \in E_Q^{\pi_k}$ if $Y_{iq}^{\pi_k} = 1$.

@@

\begin{equation} \label{eq:scores}
	\tilde{s} = \sum_{k=0}^{K} \sum_{i=1}^{m} \sum_{j=1}^{n^{\pi_k}} X_{ij}^{\pi_k} \times t_j^{\pi_k}.s
\end{equation}

\begin{equation} \label{eq:costs}
	\tilde{c} = c_{0,\pi(1)}\times(1+c_g\times|\epsilon_0|) + \sum_{k=1}^{K-1} [ c^{\pi(k), \pi(k+1)}\times(1+c_g\times|\epsilon^{\pi_k}|) ] + c^{\pi(K),0}\times(1+c_g\times|\epsilon_K|)
\end{equation}

@@

\begin{equation} \label{eq:tau}
\tau^{\pi_k} = \sum_{i=1}^{m}[ p_i.d \times (\sum_{j=1}^{n^{\pi_k}} X_{ij}^{\pi_k} \times t_j^{\pi_k}.w +  \sum_{q=1}^{m} Y_{iq}^{\pi_k} \times a_q^{\pi_k}.w)];\ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:eps}
\epsilon^{\pi_k} = \frac{\tau^{\pi_k}}{maxW \times limit^{CG}_{long}};\ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp11}
	L_0 = L; L^{\pi_k} = L^{\pi(k-1)} - \{l^{\pi_k}\}; \ k \in \{1, 2, \ldots, K\}
\end{equation}

@@

\begin{equation} \label{eq:pdp12}
	X_{ij}^{\pi_k} = 0 \ \mbox{if} \ t_j^{\pi_k}.to \notin L^{\pi_k}; \ i \in \{1, 2, \ldots, m\}; \ j \in \{1, 2, \ldots, n^{\pi_k}\}; \ k \in \{1,2, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp9}
	Y_{iq}^{\pi_k} = 0 \ \mbox{if} \ a_i^{\pi_k}.to \notin L^{\pi_k}; \ i \in \{1, 2, \ldots, m\}; \ q \in \{1, 2, \ldots, m\}; \ k \in \{ 1, 2, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:cons2}
	a_i^{\pi(k+1)}.w = \sum_{j=1}^{n^{\pi_k}} X_{ij}^{\pi_k} \times t_j^{\pi_k}.w + \sum_{q=1}^{m} Y_{iq}^{\pi_k} \times a_q^{\pi_k}.w;  \ i \in \{1, 2, \ldots, m\}; k \in \{0, 1, \ldots, K-1\}
\end{equation}

\begin{equation} \label{eq:cons3}
	a_i^{\pi(k+1)}.v = \sum_{j=1}^{n^{\pi_k}} X_{ij}^{\pi_k} \times t_j^{\pi_k}.v + \sum_{q=1}^{m_k} Y_{iq}^{\pi_k} \times a_q^{\pi_k}.v;  \ i \in \{1, 2, \ldots, m\}; k \in \{0, 1, \ldots, K-1\}
\end{equation}

\begin{equation} \label{eq:cons5}
	a_i^{\pi(k+1)}.to = t^{\pi_k}_j.to  \ \mbox{if} \ X^{\pi_k}_{ij} = 1  \ \mbox{and} \ t^{\pi_k}_j.to \in L_{\pi(k+1)};  \ i \in \{1, 2, \ldots, m\}; \ j \in \{1, 2, \ldots, n^{\pi_k}\}; \ k \in \{0,1, \ldots, K-1\}
\end{equation}

\begin{equation} \label{eq:LatIt}
	LatIt^{\pi_k} = \sum_{i=1}^{m} \sum_{j=1}^{n^{\pi_k}} ( X_{ij}^{\pi_k} \times t_j^{\pi_k}.w \times (i\%2) - X_{ij}^{\pi_k} \times t_j^{\pi_k}.w \times (i+1)\%2 ); \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:LatCons}
	LatCons^{\pi_k} =  \sum_{i=1}^{m} \sum_{q=1}^{m}  ( Y_{iq}^{\pi_k} \times a_q^{\pi_k}.w \times (i\%2) - Y_{iq}^{\pi_k} \times a_q^{\pi_k}.w \times (i+1)\%2); \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:torqlat}
	s.t.: d^{CG}_{pallet} \times | LatIt^{\pi_k} + LatCons^{\pi_k} | \leq  \sum_{i=1}^{m}p_i.w \times limit^{CG}_{lat}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:torqlong}
	s.t.: |\tau^{\pi_k}| \leq maxW \times limit^{CG}_{long};\ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:payload}
	s.t.: \sum_{i=1}^{m} (\sum_{j=1}^{n^{\pi_k}} X_{ij}^{\pi_k} \times t_j^{\pi_k}.w + \sum_{q=1}^{m_k} Y_{iq}^{\pi_k} \times a_q^k.w ) \leq maxW; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app2}
	s.t.: \sum_{j=1}^{n^{\pi_k}} X_{ij}^{\pi_k} \times t_j^{\pi_k}.w + \sum_{q=1}^{m_k} Y_{iq}^{\pi_k} \times a_q^{\pi_k}.w  \leq p_i.w; \ i \in \{1, 2, \ldots, m\}; \ {\pi_k} \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app3}
	s.t.: \sum_{j=1}^{n^{\pi_k}} X_{ij}^{\pi_k} \times t_j^{\pi_k}.v + \sum_{q=1}^{m} Y_{iq}^{\pi_k} \times a_q^{\pi_k}.v  \leq\ p_i.v; \ i \in \{1, 2, \ldots, m\}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app4}
	s.t.: \sum_{i=1}^{m} X_{ij}^{\pi_k} \leq 1; \ j \in \{1, 2, \ldots, n^{\pi_k}\}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:app5}
	s.t.:  Y_{iq}^{\pi_k} = 1 \ \mbox{if} \ a^{\pi_k}_q.to \in L^{\pi_k}; \ q \in \{1, 2, \ldots, m\}; \ k \in \{0, 1, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp8}
	s.t.: p_i.to^{\pi_k} = t^{\pi_k}_j.to\ \mbox{if} \ X_{ij}^{\pi_k} = 1; \ i \in \{1, 2, \ldots, m\}; \ j \in \{1, 2, \ldots, n^{\pi_k}\}; \ k \in \{1, 2, \ldots, K\}
\end{equation}

\begin{equation} \label{eq:pdp2}
	s.t.:  p_i.to^{\pi_k} = a^{\pi_k}_q.to\ \mbox{if} \ Y_{iq}^{\pi_k} = 1; \ i \in \{1, 2, \ldots, m\};\ q \in \{1, 2, \ldots, m\}; \ k \in \{1, 2, \ldots, K\}
\end{equation}

The objective of this problem is to find a permutation $\pi \in S_K$\/ that maximizes the function $f_\pi(\tilde{s},\tilde{c})$ \ref{eq:maxf}. In this way, the flight plan will be $l_0, l_{\pi(1)}, \ldots, l^{\pi_k}, l_0$. $\tilde{s}$\/ is the total score of transported items \ref{eq:scores} and  $\tilde{c}$\/ is the total cost of fuel consumed \ref{eq:costs}. As can be seen, $\tilde{c}$\/ corresponds to the fuel consumption due to the flights carried out and the CG deviation of the transported cargo. Throughout this work, for simplicity, we use $f=\tilde{s}/\tilde{c}$.

@@

The tactic will be, at each shipping node, to predefine the destinations of the pallets at that node. In this way, we will reserve a number of pallets proportional to the volume demanded by each destination at the shipping node. We could have used another criterion, but it was observed in the experiments that volume is more constrictive in airlift. Once the destinations of the pallets are defined, we will use serial and multiprocessing heuristics to find the best possible node-by-node solutions. This strategy is summarized in Algorithm \ref{alg:main}, retrieved from \cite{MesquitaSanches2023}.

@@

The other parameter $volume$\/ is a value greater than 1, which corresponds, at each node $\pi_k$, to the ratio between the sum of the volumes of the items ($\sum_{j=1}^{n^{\pi_k}} t^{\pi_k}_j.v$) and the load capacity of the pallets ($\sum_{i=1}^{m} p_i.v$). This parameter is passed to $ItemsGeneration$ (line \ref{main:items}), responsible for creating the items to be shipped, which will be presented in the next section.

@@

As we mentioned in the previous section, all tours start and end at $\pi_0$\/ (lines \ref{tour:pi1}-\ref{tour:pi2}).
After initializing the score and cost values (lines \ref{tour:score}-\ref{tour:cost}), there is a loop for the $K+1$\/ flights (lines \ref{tour:loop1}-\ref{tour:loop2}). Initially we set pallets destination as $-1$\/ (line \ref{tour:-1}). When the aircraft is at node $\pi_0$, the initial graph $G_1$\/ is empty because it has no consolidated items \ref{tour:g11}. Otherwise, the set $L_k$\/ of remaining nodes when departing from $\pi_k$ is updated (line \ref{tour:lk1}), and $UpdateConsolidated$\/ (line \ref{tour:dest}) returns the set of consolidated items that have not yet reached their destinations and remain on board, rearranging them on the pallets to minimize CG deviation. This allocation is stored in graph $G_1$\/ (line \ref{tour:g12}).

@@

		\State $\pi[0] \gets 0$ \label{tour:pi1}
		\State $\pi(K+1) \gets 0$ \label{tour:pi2}
		\State $score \gets 0$ \label{tour:score}
		\State $cost \gets 0$ \label{tour:cost}
		\For {$k\gets0$ to $K$} \label{tour:loop1}		
			\For{$i \gets 1$ to $m$}
				\State $p_i.to[k] \gets -1$ \label{tour:-1} \Comment{reset this pallet destination}
			\EndFor	
			\If {$k = 0$}
				\State $L_0 \gets L$
				\State Let $G_1(M \cup N_0, \varnothing)$ \label{tour:g11}
			\Else
				\State $L_k \gets L_k - \pi(k)$  \label{tour:lk1}			
				\State $Q^{\pi_k}, E_Q^{\pi_k}, M^{\pi_k} \gets UpdateConsolidated(\pi(k))$ \label{tour:dest}			
				\State Let $G_1(M^{\pi_k} \cup N^{\pi_k} \cup Q^{\pi_k}, E_Q^{\pi_k})$ \label{tour:g12}
			\EndIf  \label{tour:lk2}	
			\State $M^{\pi_k} \gets SetPalletsDestination( \pi(k) )$ \label{tour:dest2}		
			\State $G_2 \gets SolveNode(method, \pi(k), G_1, timeLim)$ \label{tour:node}

@@

Then $MinCGDeviation$ is run through a MIP solver to relocate the consolidated items on the pallets minimizing torque and ensuring that they all remain on board, one on each pallet. As there are few variables, the MIP solver returns an allocation $E_Q^{\pi_k}$ very quickly.

@@

In this subsection we present implementations of $SolveNode(...)$, where $method$\/ corresponds to a heuristic, $k$\/ is the index of the current node $\pi_k$ and $G$\/ is the allocation graph of the consolidated items that remain on board at $\pi_k$.

@@

The selection of edges for $E_N^{\pi_k}$\/ uses the {\it edge attractiveness}\/ $\theta^{\pi_k}_{ij}$, Equation \ref{eq:edge}, which can be understood as the tendency to allocate the item $t^{\pi_k}_j$\/ to the pallet $p_i$. It is directly proportional to the score, and inversely relative to the volume and torque of each item.

\begin{equation} \label{eq:edge}
	\theta^{\pi_k}_{ij}= \frac{t^{\pi_k}_j.s}{t^{\pi_k}_j.v \times 3000}\times(1-\frac{t^{\pi_k}_j.w\times|p_i.d|}{\max_s\{t^{\pi_k}_s.w\}\times\max_q\{|p_q.d|\}});\ i \in \{1,2,\ldots,m\},\ j \in \{1,2,\ldots,n_k\}
\end{equation}

@@

In the pseudo-codes presented below, given an allocation graph $G$\/ for the node $\pi_k$, $f_s(G)$\/ is equivalent to the score $s$\/ returned by $ScoreAndDeviation({\pi_k},G) $, described in \cite{MesquitaSanches2023}.

@@

According to \cite[p.226]{manfrin2006}, there are five topologies for multiprocessing interchange of information: {\it fully-connected}, where the master process broadcasts to all remaining child processes; {\it replace-worst}, where the best-so-far solution process broadcasts only to the current worst solution process; {\it hypercube}, where processes are connected as a hypercube, and a vertex process broadcasts only to the connected vertices; {\it ring}, in which one process only sends a message to the next process connected to it; and {\it parallel independent runs (PIR)}, in which there are no communication costs and the best solution is chosen among all processes. They concluded that the PIR topology presented, in average, the smallest distance from the optimal solutions.

@@

$3^{st}$ - {\it Shims} set construction and best {\it Shims} selection
ess run-time, we considered each pallet as a parallel process, taking advantage of nowadays multi-core computers.

@@

		\Procedure{$mpShims$}{${N^{\pi_k}, \pi(k)}, limit, G$}

		\State Let $G(M^{\pi_k},\ Q^{\pi_k},\ E^{\pi_k})$ \label{mpShims:g1} \Comment{the initial solution}
